---
title: "Week of 2/9/2020 Assignment"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

#### Assignment: Tidyverse Review of Data

*Make sure all code is written in Tidyverse.
1. Download the Titanic Dataset
    + Download both test and train csv files.
    + https://github.com/syang1984/AnalyticsAssignments
2. Load the train dataset into your workspace.
3. Review train dataset.
    + Provide summary of data for dataset.
      + What conclusions can be discerned from the dataset?
4. Present the data in one of the following comparing various variables to survival:
    + Bar Graph
    + Box plot
    + Bar graph
    + Scatter plot
5. Ensure graphs are color coded, labeled and makes sense.
    + Export graphs and post to github.
6. Provide short narrative conclusion in RMD.
7. Label files as such: 20200209_Assignment_(name)_(initials)
  
#### Code 
```{r setup, include=TRUE, echo=TRUE}
library(tidyverse)
```
# Load train dataset
```{r Load Data}
setwd("C:/Users/labmo/Documents/R/R projects/Titanic R Project/")
train <- read_csv("titanic_train.csv.txt")
```
```{r Survial Logic}
train_log <- mutate(train, Survived = as.logical(train$Survived))
```
Provide a summary of data for dataset train
```{r Sum Up}
train_log %>% 
  summary()
```

From the brief summary, we see that in this dataset: passenger IDs range from 1 to 891 (a total of 891 observations); Passenger Class goes from 1 to 3; The youngest passenger was under 6 months old; The oldest passenger was 80 years old; The mean passenger age was 29.7 years; The minimum fare was listed at $0.00; Most fares ranged from $7.91 to $31.00, with the mean fare paid being $32.20; The maximum fare listed was $512.33.


They say women and children first, but which sex has more survivors in this data set?
```{r Bar Plot}
train_log %>% 
  filter(Survived == 1) %>% 
ggplot(aes(x = Sex)) +
  geom_bar(fill = "steelblue4") +
  labs(x = "Passenger Sex", y = "Number of Survivors", title = "The Gender Gap: Survival at Sea")
```


The number of female survivors far outpaces the number of male survivors here.


What role did Class and Fare price play in survivorship?
```{r Point Plot}
train_rmna <- train_log %>% 
  filter(Pclass != "NA", Fare != "NA", Survived != "NA")
ggplot(train_log, aes(x = Pclass, y = Fare, color = Survived))+
  geom_point(position = "jitter", na.rm = TRUE)+
  scale_x_discrete(limits= c(1, 2, 3))+
  scale_y_sqrt()+
  labs(x = "Passenger Class", y = "Fare Paid", title = "The Role of Wealth & Class in Survivorship")
```


As expected, the most expensive fares were for 1st class.  As for survival, 1st and 2nd class passengers had more survivors in their groups, and a trend toward more survivors the higher ther fare within those classes.  For passengers in 3rd class, lower fares came with lower survival.

But the real question is- won't somebody think of the children?
```{r Box Plot}
train_rmna %>% 
  ggplot(aes(x=Survived, y=Age))+
  geom_boxplot(aes(fill = Survived, color = Survived))+
  scale_fill_manual(values = c("seagreen4", "skyblue3")) +
  scale_color_manual(values = c("black", "black")) +
  labs(x = "Survived T/F", y = "Passenger Age", 
       title = "The Affect of Age on Survivorship")
```


Overall, age of survivors was slightly younger than that of non-surviors, though the mean age of both survivors and non-survivors were nearly identical.

In this data set, it appears being wealthy, young, and female was a good bet to survive.