---
title: "Secret Santa Data Analysis Project"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
{eval=TRUE}
# Make sure all code is written in Tidyverse.
library(tidyverse)


## 1. Read submitted csv from fellow classmates.
## 2. Load data into your own work space.

# Classmate's data set saved as "set_a"
set_a <- read_csv("C:/Users/labmo/Documents/R/R projects/Santa Project/Assignment_20200119_SY.txt")
View(set_a)
# My dataset saved as "set_b"
set_b <- read_csv("C:/Users/labmo/Documents/R/R projects/Santa Project/ExchangeMatches.csv")
View(set_b)

# Renaming columns so set_a and set_b use the same terms
abnames <- c("Giver", "Recipient", "Gift_Price", "Purchase_Date")
colnames(set_a) <- abnames
colnames(set_b) <- abnames
```

In this document, we'll be analyzing the gift-buying activity of two population groups, which are designated "Set A" and "Set B".  We will start by viewing the summary of the data from Set A.
```{r, echo=FALSE}
```

<center> __Set A Summary:__ </center>
```{r}
sumup_a <- set_a %>% 
  summary()
sumup_a
```

Looking at the data for Set A, the majority of participants spent between $36.01 and $82.62 on their gift.  The cheapest gift purchased was $6.41, while the most expensive gift was purchased for $364.97.  When compared to our median, it appears this maximum gift price may be out of the ordinary for the spending of Set A.

Most gifts purchased in Set A were bought between December 1st and December 3rd 2019, which point towards early-bird spending habits.  However, the latest date of purchase was still the last day a gift could be purchased (December 5th 2019), meaning that even though fewer gift purchases happened later, they did occur across the range of all possible dates.

The earliest purchase date was April 1st 2019, and at a glance it apppears this date may be an outlier as well- we'll examine this further in later analysis.

Next, we'll examine the summary of Set B

<center> __Set B Summary:__ </center>
```{r}
sumup_b <- set_b %>% 
  summary()
sumup_b
```

In this population, we see the most participants spent between $31.14 and $74.24 on their gifts, with the minimum price paid being $9.64.  The most expensive gift in Set B was $364.97, the same as in Set A.  Again, with this price being over 6.5 times the median gift price, this may be outside the typical spending for this population as well.

Examining the purchase dates reveals that in Set B most gifts were purchased between December 1st and December 3rd 2019, with a trend toward earlier shopping that matches Set A's habits.  The latest purchase date for Set B is December 4th, which shows an earlier end to this population's gift buying when compared to Set A.  Once again we see a likely outlier in the earliest purchase date for Set B- April 1st 2019.

<center> __Comparing Spending Habits: Cheapskates__ </center>

Doing some quick comparisons of the summaries, Set B has a lower mean, median, and 1st quartile than Set A- we may assume that Set B then has a higher number of participants prefering to buy cheaper gifts (<$20).  But is this the case?
```{r}
cheap_a <- c(set_a$Gift_Price < 20) %>% 
  grep(pattern = TRUE)
cheap_a

cheap_b <- c(set_b$Gift_Price < 20) %>%
  grep(pattern = TRUE)
cheap_b
```

Yes, as we can see from this comparison that Set B has 3 cheapskates, while Set A only has 1 cheapskate this year.

<center> __Comparing Shopping Habits: Procrastinators__ </center>

When comparing purchase dates, in both Set A and Set B, 25% of participants had purchased their gift on  December 1st 2019. However the latest purchase date for Set A was 2019-12-05, while the lastest purchase date for Set B was December 4th 2019.  Does this mean participants in Set A were more likely to procrastinate in their purchasing than those in Set B? 

```{r}
proc_a <- c(set_a$Purchase_Date >= "2019-12-04") %>% 
  grep(pattern = TRUE)
proc_a

proc_b <- c(set_b$Purchase_Date >= "2019-12-04") %>%
  grep(pattern = TRUE)
proc_b
```
No- in fact 4 participants in Set B waited until at least December 4th 2019 to make their gift purchase, vs only 3 participants in Set A waiting until at least December 4th to buy their gift.

Looking at the data sets, we can see Chris spent a lot on his gift for Carla, and bought this gift very early in comparison to the rest of the participants in each data set. 

<center> __View Set A:__ </center>

```{r}
print(set_a)
```

<center> __View Set B:__ </center>
```{r}
print(set_b)
```

What does the summary of our data look like for our average participant, filtering Chris's activity out of the group (as an outlier)?

<center> __Set A (filtered):__ </center>
```{r}
set_axc <- filter(set_a, Giver != "Chris")
set_axc %>% 
  summary()
```
<center> __Set B (filtered):__ </center>
```{r}
set_bxc <- filter(set_b, Giver != "Chris")
set_bxc %>% 
  summary()
```
We now see that the mean is much lower for both groups, and a new max can be seen.  In (filtered) Set A, the mean amount spent on a gift was $58.99, with most participants spending more than $30.00 and less than $79 on a gift.  The most expensive gift was purchased for $98.61.

In (filtered) Set B, the mean amount spent on a gift was $48.87, with most participants spending at least $27.00, but less than $73.00 on their purchased gift.  The most expensive gift in (filtered) set_b was purchased for $88.81.

We could then expect to sell slightly higher cost gifts on average in stores frequented by participants in Set A vs those in Set B.

<center> __Visual Interpretations__ </center>

Let's get down to the visuals.

I added some additional research I've gathered to the data sets to further compare these sets and also evaluate each sets by family group.
```{r}
# Create a variable to group/identify which set observations belong to
set_a <- set_a %>% 
  mutate(Set = c("Set_A")) %>% 
  mutate(Family = c("Scrubs", "Scrubs", "Boat", "Boat", "Parks", "Parks", 
                    "Modern", "Modern", "Office", "Office", "Office",
                    "Parks", "Boat", "Parks"))

set_b <- set_b %>% 
  mutate(Set = "Set_B") %>% 
  mutate(Family = c("Scrubs", "Scrubs", "Office", "Office", "Parks",
                    "Parks", "Modern", "Modern", "Boat", "Boat",
                    "Office", "Parks", "Boat", "Parks"))
```

How does spending compare amongst the Families in each Set?

<center>
```{r}
##Pie Chart A
# Prepare data to compare which Family group in each set spent the most overall on gifts
spend_a <- set_a %>% 
  group_by(Family) %>% 
  summarise(Total_Spent = sum(Gift_Price))
pie_label_a <- paste(spend_a$Family, " ", "$", spend_a$Total_Spent)
pie_set_a = pie(spend_a$Total_Spent, labels = pie_label_a, edges = 300, main = "Total Spending by Family: Set A")
```
```{r}
spend_b <- set_b %>% 
  group_by(Family) %>% 
  summarise(Total_Spent = sum(Gift_Price))
pie_label_b <- paste(spend_b$Family, " ", "$", spend_b$Total_Spent)

pie_set_b = pie(spend_b$Total_Spent, labels = pie_label_b, edges = 300, main = "Total Spending by Family: Set B")
```
</center>
Chris's large gift purchase price puts his family (Scrubs) in the highest total spending bracket for both Set A and Set B.  

In Set A, we see the Parks family has the second highest total spent, followed by Office & Modern, with Boat family having the least total spent.

In Set B, the Parks family again has the second highest total spent, however now Boat & MOdern have the next highest totals, with Office family having the lowest total spent.

Some of these difference may be influenced by family size (the number of participants from each family in this exchange), though this is not witness evenly across Sets (ex: Modern family having only two participants outspending families with 3 participants in each Set).

```{r, echo=FALSE}
# Create combination sets to view a side-by-side comparison in charts
set_combo <- set_a %>% 
  bind_rows(set_b)


set_axc <- set_axc %>% 
  mutate(Set = c("Set_A"))
set_bxc <- set_bxc %>% 
  mutate(Set = "Set_B")
set_comboxc <- set_axc %>% 
  bind_rows(set_bxc)
```
Let's take a look at the purchase dates of each group (excluding Chris's outlier) in a visual form.

<center>
```{r}
ggplot(set_comboxc, aes(x = Purchase_Date)) +
  geom_bar(aes(fill = Set)) +
  scale_fill_manual(values = c("Set_A" = "dodgerblue2", "Set_B" = "firebrick")) +
  facet_wrap(~Set) +
  labs(x = "Date of Purchase", y = "Number of Gifts Purchased", title = "Gifts Purchased by Date: Comparing Sets")+
  theme(plot.title = element_text(hjust = 0.5))
```
</center>

As seen in the bar graphs, December 1st 2019 was the most popular day to purchase gifts for participants in Set A, while those in set B were more likely to shop mid-week, prefering December 3rd and December 4th 2019 for making their purchases. 


Lastly, we'll see our summarized data in a boxplot.
<center>
```{r}
ggplot(set_combo, aes(x = Set, y = Gift_Price)) +
 geom_boxplot(aes(fill = Set, color = Set)) +
  scale_fill_manual(values = c("dodgerblue2", "firebrick")) +
  scale_color_manual(values = c("black", "black")) +
  labs(x = "Population/Set", y = "Price of Gift Purchased", 
       title = "Comparison of Spending on Gifts Between Sets") +
  scale_y_log10() +
  theme(plot.title = element_text(hjust = 0.5))
```
</center>
The boxplots visually confirm what was suspected earlier in viewing the set summaries- Chris's expensive gift purchase is an outlier in each set.  We can also see that the spending in Set A was much more tightly grouped than in Set B; Set A participants (outliers excluded), spent between ~$30 to ~$100 on their gifts.  The spending from Set B was much more spread out, rangding from ~$10 to ~$90 on gift purchases.  Average spending on gifts was lower in Set B than in Set A, though Set A did have the lowest price paid for a gift purchased.

It appears from this data that gift shops frequented by the population in Set A can expect to sell slightly higher priced items earlier in the shopping week.  Shops frequented by the population in Set B may benefit from stocking slightly lower priced gifts and expect more purchases to take place mid-to-late week.  However, both sample sets are small, and the differences may not be statistically significant.